Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-10-27 WSSSPE3/report/WSSSPE3_review.tex
working on conclusions

\todo{need to update this section for WSSSPE3}

The WSSSPE3 workshop continued our experiment from WSSSPE1 and WSSSPE2 in how we can
collaboratively build a workshop agenda, and we began a new experiment in
how to build a series of workshops into an ongoing community activity.

The differences in workshop organization in WSSSPE2 from WSSSPE1
are in using an existing service (EasyChair) to handle submissions and reviews,
rather than an ad hoc process, and using an existing service (Well Sorted) to
allow collaborative grouping of papers into themes by all authors, reviewers,
and the community, rather than this being done in an ad hoc manner by the
organizers alone.

The fact remains that contributors also want to get credit for their
participation in the process. And the workshop organizers will want to make
sure that the workshop content and their efforts are recorded. Ideally, there
would be a service that would index the contributions to the
workshop, serving the authors, the organizers, and the larger community. 
Since there still isn't such a service today, the workshop organizers are
writing this initial report and making use of arXiv as a partial solution to
provide a record of the workshop.

@@

In terms of building community activities, we wanted to focus primarily on
working groups, which we were able to do, as discussed above, but we
also wanted to make sure that attendees felt they had a chance to get their
ideas across to the whole group, which was the purpose of the lightning talks.
Overall, this seemed to be successful at the time, in terms of both the lightning
talks and the breakout groups, and the discussion of sustainability also led
to interesting and useful results. However, the challenge that we have discovered
since WSSSPE2 is that it is very hard to continue the breakout groups'
activities.  The WSSSPE2 participants were willing to dedicate their time to
the groups while they were at the meeting, but afterwards, they have gone
back to their (paid) jobs.  We need to determine how to tie the WSSSPE
breakout activities to people's jobs, so that they feel that continuing them
is a higher priority than it is now, perhaps through funding the participants,
or through funding coordinators for each activity, or perhaps by getting
the workshop participants to agree to a specific schedule of activities during the
workshop.

>>>>

In WSSSPE3, we attempted to take what we learned from WSSSPE1 and WSSSPE2
in how we can collaboratively build a workshop agenda and turn that into
an ongoing community activity.  The success or failure of this effort
of this will only become apparent over time.

The workshop had two components, presentations and working groups.  The presentations,
in the first half day of the workshop, including a inspirational keynote and a set of lightning
talks.
We used lightning talks for two reasons: first, the need of some participants to have a slot
on the agenda to justify their attendance, and second, as a way to get new ideas across
to all the attendees.  We broke with the tradition of requiring the lightning talk submitters
to self-publish their papers, and instead used a common peer-review platform\footnote{\url{http://easychair.org}}, choosing to publish the speaker slides on the workshop web site
instead.

The working groups met for a small part of the first half day and all of the second day,
excepting some times for the groups to report back to the collected workshop attendees.
\todo{need to say some more about what came of this}

\todo{this paragraph is from the wssspe2 report - to be updated...}
However, the challenge that we have discovered
since WSSSPE2 is that it is very hard to continue the breakout groups'
activities.  The WSSSPE2 participants were willing to dedicate their time to
the groups while they were at the meeting, but afterwards, they have gone
back to their (paid) jobs.  We need to determine how to tie the WSSSPE
breakout activities to people's jobs, so that they feel that continuing them
is a higher priority than it is now, perhaps through funding the participants,
or through funding coordinators for each activity, or perhaps by getting
the workshop participants to agree to a specific schedule of activities during the
workshop.

@@
